---
title: "Task Completion Model"
author: Hampus Broman & William Levén
date: 2021-05
output: 
  html_document: 
    pandoc_args: [ "-o", "html/task_completion.html" ]
---

```{r include-setup, include=FALSE}
# Load setup file
source(knitr::purl('setup.Rmd', output = tempfile()))
```

## Looking at the data {.tabset}

We plot the data and can see that there is no obvious large difference between the debt versions. We do however see a difference when it comes to which scenario the dropout was on.

### Per session

```{r plo1}
d.sessions %>%
  ggplot(aes(task_completion, fill = task_completion)) +
  geom_bar() +
  labs(title = "Task completion per session") +
  xlab("Least completed task per session") +
  ylab("Number of sessions") +
  scale_fill_manual(values = rep("#7070FF", 4), guide = NULL)
```

### Per debt level

```{r plot2}
d %>%
  mutate_at("high_debt_version", 
            function(x) case_when(
              x == "false" ~ "Low debt",
              x == "true" ~ "High debt"
            )) %>%
  filter(task_completion == "Not submitted") %>%
  ggplot(aes(x=high_debt_version, fill = high_debt_version)) +
  geom_bar() +
  labs(title = "Droputs per debt level") +
  xlab("Debt level") +
  ylab("Number of dropouts") +
  scale_fill_manual(values = rep("#7070FF", 2), guide = NULL) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8))
```

### Per scenario

```{r plot3}

d %>%
  filter(task_completion == "Not submitted") %>%
  ggplot(aes(scenario, fill = scenario)) +
  geom_bar() +
  labs(title = "Dropputs per scenraio") +
  ylab("Number of dropouts") +
  xlab("Scenario") +
  scale_fill_manual(values = rep("#7070FF", 2), guide = NULL) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10))
```

### Per signup group

```{r plot4}
d.sessions %>%
  mutate_at("group", 
            function(x) x <- case_when(
              x == "code-interested" ~ "code-int.",
              x == "product-company" ~ "prod.-comp.",
              x == "professional-contact" ~ "prof.-cont.",
              TRUE ~ as.character(x)
            )) %>%
  ggplot(aes(group, fill=task_completion, position = "fill")) +
  geom_bar(position = "fill") +
  labs(title = "Dropputs per signup group") +
  ylab("Ratio") +
  xlab("Signup Group") +
  scale_fill_manual("Task completion", values = c("transparent", "lightblue", "#7070FF", "darkblue"))
```

## Initial model
The outcome model represent incremental steps and will therefore be modeled with stopping ratio

We include `high_debt_verison` as a predictor in our model as this variable represent the very effect we want to measure.

### Selecting Priors {.tabset}

We iterate over the model until we have sane priors.

#### Base model with priors

```{r initial-model-definition, class.source = 'fold-show'}
dropouts.with <- extendable_model(
  base_name = "dropouts",
  base_formula = "task_completion ~ 1 + high_debt_version",
  base_priors = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = d,
)
```


#### Default priors

```{r default-priors}
prior_summary(dropouts.with(only_priors= TRUE))
```

#### Selected priors

```{r selected-priors}
prior_summary(dropouts.with(sample_prior = "only"))
```

#### Prior Predictive Check

```{r priors-check, warning=FALSE}
pp_check(dropouts.with(sample_prior = "only"), nsamples = 200, type = "bars")
```

#### Beta Parameter Influence

```{r priors-beta, warning=FALSE}
sim.size <- 1000
sim.intercept <- rnorm(sim.size, -2, 1)
sim.beta <- rnorm(sim.size, 0, 0.4)
sim.beta.diff <- (plogis(sim.intercept + sim.beta) / plogis(sim.intercept) * 100) - 100

data.frame(x = sim.beta.diff) %>%
  ggplot(aes(x)) +
  geom_density() +
  xlim(-80, 80) +
  labs(
    title = "Beta parameter prior influence",
    x = "Estimate with beta as % of estimate without beta",
    y = "Density"
  )

```

### Model fit {.tabset}

We check the posterior distribution and can see that the model seems to have been able to fit the data well. 
Sampling seems to also have worked well as Rhat values are close to 1 and the sampling plots look nice.

#### Posterior Predictive check

```{r base-pp-check}
pp_check(dropouts.with(), nsamples = 200, type = "bars")
```

#### Summary

```{r base-summary}
summary(dropouts.with())
```

#### Sampling plots

```{r base-plot}
plot(dropouts.with(), ask = FALSE)
```

## Model predictor extenstions {.tabset}

```{r mo-priors}
# default prior for monotonic predictor
edlvl_prior <- prior(dirichlet(2), class = "simo", coef = "moeducation_level1")
```

We use `loo` to check some possible extensions on the model.

### One variable {.tabset}

```{r model-extension-1, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  dropouts.with(),
  
  # New model(s)
  dropouts.with("work_domain"),
  dropouts.with("work_experience_programming.s"),
  dropouts.with("work_experience_java.s"),
  dropouts.with("education_field"),
  dropouts.with("mo(education_level)", edlvl_prior),
  dropouts.with("workplace_peer_review"),
  dropouts.with("workplace_td_tracking"),
  dropouts.with("workplace_pair_programming"),
  dropouts.with("workplace_coding_standards"),
  dropouts.with("scenario"),
  dropouts.with("group")
)
```

#### Comparison

```{r model-extension-1-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-1-dig, warning=FALSE}
loo_result[1]
```

### Two variables {.tabset}

```{r model-extension-2, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  dropouts.with(),
  
  dropouts.with("education_field"),
  dropouts.with("workplace_pair_programming"),
  dropouts.with("scenario"),
  dropouts.with("group"),
  
  # New model(s)
  dropouts.with(c("education_field", "workplace_pair_programming")),
  dropouts.with(c("education_field", "scenario")),
  dropouts.with(c("education_field", "group")),
  
  dropouts.with(c("workplace_pair_programming", "scenario")),
  dropouts.with(c("workplace_pair_programming", "group")),
  
  dropouts.with(c("scenario", "group"))
)
```

#### Comparison

```{r model-extension-2-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-2-dig, warning=FALSE}
loo_result[1]
```

### Three variables {.tabset}

```{r model-extension-3, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  dropouts.with(),
  
  dropouts.with("education_field"),
  dropouts.with("workplace_pair_programming"),
  dropouts.with("scenario"),
  dropouts.with("group"),
  
  dropouts.with(c("education_field", "workplace_pair_programming")),
  dropouts.with(c("education_field", "scenario")),
  dropouts.with(c("workplace_pair_programming", "scenario")),
  dropouts.with(c("education_field", "group")),
  
  # New model(s)
  dropouts.with(c("education_field", "workplace_pair_programming", "scenario")),
  dropouts.with(c("group", "workplace_pair_programming", "scenario")),
  dropouts.with(c("education_field", "group", "scenario")),
  dropouts.with(c("education_field", "workplace_pair_programming", "group"))
)
```

#### Comparison

```{r model-extension-3-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-3-dig, warning=FALSE}
loo_result[1]
```

### Four variables {.tabset}

```{r model-extension-4, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  dropouts.with(),
  
  dropouts.with("education_field"),
  dropouts.with("workplace_pair_programming"),
  dropouts.with("scenario"),
  dropouts.with("group"),
  
  dropouts.with(c("education_field", "workplace_pair_programming")),
  dropouts.with(c("education_field", "scenario")),
  dropouts.with(c("workplace_pair_programming", "scenario")),
  dropouts.with(c("education_field", "group")),
  
  dropouts.with(c("education_field", "workplace_pair_programming", "scenario")),
  dropouts.with(c("education_field", "workplace_pair_programming", "group")),
  
  # New model(s)
  dropouts.with(c("education_field", "workplace_pair_programming", "group", "scenario"))
)
```

#### Comparison

```{r model-extension-4-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-4-dig, warning=FALSE}
loo_result[1]
```


## Candidate models  {.tabset}
We pick some of our top performing models as candidates and inspect them closer.

The candidate models are named and listed in order of complexity.

### Dropouts0  {.tabset}

We select the simplest model as a baseline.

```{r droputs0, class.source = 'fold-show'}
droputs0 <- brm(
  "task_completion ~ 1 + high_debt_version",
  prior = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = as.data.frame(d),
  file = "fits/droputs0",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r droputs0-sum}
summary(droputs0)
```

#### Sampling plots

```{r droputs0-plot}
plot(droputs0, ask = FALSE)
```

#### Posterior predictive check

```{r droputs0-pp}
pp_check(droputs0, nsamples = 200, type = "bars")
```

### Dropouts1  {.tabset}

We select the best performing model with one variable.

```{r droputs1, class.source = 'fold-show'}
droputs1 <- brm(
  "task_completion ~ 1 + high_debt_version + education_field",
  prior = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = as.data.frame(d),
  file = "fits/droputs1",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r droputs1-sum}
summary(droputs1)
```

#### Sampling plots

```{r droputs1-plot}
plot(droputs1, ask = FALSE)
```

#### Posterior predictive check

```{r droputs1-pp}
pp_check(droputs1, nsamples = 200, type = "bars")
```

### Dropouts2  {.tabset}

We select the best performing model with two variables.

```{r droputs2, class.source = 'fold-show'}
droputs2 <- brm(
  "task_completion ~ 1 + high_debt_version + education_field + workplace_pair_programming",
  prior = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = as.data.frame(d),
  file = "fits/droputs2",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r droputs2-sum}
summary(droputs2)
```

#### Sampling plots

```{r droputs2-plot}
plot(droputs2, ask = FALSE)
```

#### Posterior predictive check

```{r droputs2-pp}
pp_check(droputs2, nsamples = 200, type = "bars")
```

### Dropouts3  {.tabset}

We select the best performing model with three variables.

```{r droputs3, class.source = 'fold-show'}
droputs3 <- brm(
  "task_completion ~ 1 + high_debt_version + education_field + workplace_pair_programming + scenario",
  prior = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = as.data.frame(d),
  file = "fits/droputs3",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r droputs3-sum}
summary(droputs3)
```

#### Sampling plots

```{r droputs3-plot}
plot(droputs3, ask = FALSE)
```

#### Posterior predictive check

```{r droputs3-pp}
pp_check(droputs3, nsamples = 200, type = "bars")
```

### Dropouts4  {.tabset}

We select the best performing model with four variables.

```{r droputs4, class.source = 'fold-show'}
droputs4 <- brm(
  "task_completion ~ 1 + high_debt_version + education_field + workplace_pair_programming + scenario + group",
  prior = c(
    prior(normal(0, 0.4), class = "b"),
    prior(normal(-2, 1), class = "Intercept")
  ),
  family = sratio(),
  data = as.data.frame(d),
  file = "fits/droputs4",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r droputs4-sum}
summary(droputs4)
```

#### Sampling plots

```{r droputs4-plot}
plot(droputs4, ask = FALSE)
```

#### Posterior predictive check

```{r droputs4-pp}
pp_check(droputs4, nsamples = 200, type = "bars")
```

## Final Model 
All candidate models look nice, the more complex models are slightly betters, we choose the simplest model that is not significantly worse then the best model: `dropouts2`

## Interpreting the model
To begin interpreting the model we look at how it's parameters were estimated. As our research is focused on how the outcome of the model is effected we will mainly analyze the $\beta$ parameters.

### $\beta$ parameters

```{r interpret-beta-plot, warning=FALSE, message=FALSE}
mcmc_areas(droputs2, 
           pars = c(
             "b_high_debt_versionfalse", 
             "b_education_fieldComputerScience", 
             "b_education_fieldElectricalEngineering",
             "b_education_fieldIndustrialengineering",
             "b_education_fieldInteractionDesign",
             "b_education_fieldNone",
             "b_education_fieldSoftwareEngineering",
             "b_workplace_pair_programmingfalse"
                    ),
           prob = 0.95) + scale_y_discrete() +
  scale_y_discrete(labels=c(
    "High debt version: false",
    "Education field: Computer Science",
    "Education field: Electrical Engineering",
    "Education field: Industrial Engineering",
    "Education field: Interaction Design",
    "Education field: None",
    "Education field: Software Engineering",
    "No workplace pair programming"
    )) +
  ggtitle("Beta parameters densities for task completion", subtitle = "Shaded region marks 95% of the density. Line marks the median")
```

### Effects sizes

```{r effect-size, fig.width=8}
post_settings <- expand.grid(
  high_debt_version = c("false", "true"),
  education_field = NA,
  workplace_pair_programming = NA
)

post <- posterior_predict(droputs2, newdata = post_settings) %>%
  melt(value.name = "estimate", varnames = c("sample_number", "settings_id")) %>%
  left_join(
    rowid_to_column(post_settings, var= "settings_id"),
    by = "settings_id"
  ) %>%
  select(
    estimate,
    high_debt_version
  )

post %>%
  mutate_at("high_debt_version", 
            function(x) case_when(
              x == "false" ~ "Low debt",
              x == "true" ~ "High debt"
            )) %>%
  mutate_at("estimate", 
            function(x) case_when(
              x == 1 ~ "Not submitted",
              x == 2 ~ "Does not compile",
              x == 3 ~ "Invalid solution",
              x == 4 ~ "Completed"
            )) %>%
  ggplot(aes(high_debt_version)) +
  geom_bar(aes(fill = estimate), position = "fill") + 
  scale_y_reverse() +
  scale_fill_manual("Legend", values = c("darkblue", "#7070FF", "lightblue", "transparent"), guide = guide_legend(reverse = TRUE)) +
  labs(title = "Task completion") +
  xlab("Debt version") +
  ylab("Ratio of task completion")

```

We can see that task completion ratios are very similar for both high and low debt.
