---
title: "Time To Complete Task Model"
author: Hampus Broman & William Levén
date: 2021-05
output: 
  html_document: 
    pandoc_args: [ "-o", "docs/time_to_complete_task.html" ]
---

```{r include-setup, include=FALSE}
# Load setup file
source(knitr::purl('setup.Rmd', output = tempfile()))
```


## Looking at the data

We plot the data and can see that there is no obvious large difference between the versions with high and low debt.

```{r plot}
d.both_completed %>%
  ggplot(aes(x=time/60, fill=high_debt_version)) + 
  geom_boxplot() +
  labs(
    title = "Distribution of time measurements for the different debt levels",
    subtitle = "Notice! Log10 x-scale",
    x ="Time (min)"
  ) +
  scale_y_continuous(breaks = NULL) +
  scale_x_log10() +
  scale_fill_manual(
    name = "Debt level", 
    labels = c("High debt", "Low debt"), 
    values = c("#7070FF", "lightblue"), 
    guide = guide_legend(reverse = TRUE)
  ) 
```

## Descriptive Statistics

```{r descriptive-statistics}
d.both_completed %>%
  pull(time) %>% 
  summary()

sprintf("Variance: %.0f", var(pull(d.both_completed, time)))
```

## Initial model
As the variance is much greater than the mean we will use a negative binomial family that allows us to model the variance separately.

We include `high_debt_verison` as a predictor in our model as this variable represent the very effect we want to measure.
We also include a varying intercept for each individual to prevent the model from learning too much from single participants with extreme measurements.

### Selecting Priors {.tabset}

We iterate over the model until we have sane priors.

#### Base model with priors

```{r initial-model-definition, class.source = 'fold-show'}
time.with <- extendable_model(
  base_name = "time",
  base_formula = "time ~ 1 + high_debt_version + (1 | session)",
  base_priors = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = d.both_completed,
  base_control = list(adapt_delta = 0.95)
)
```

#### Default priors

```{r default-priors}
prior_summary(time.with(only_priors= TRUE))
```

#### Selected priors

```{r selected-priors}
prior_summary(time.with(sample_prior = "only"))
```

#### Prior Predictive Check

```{r priors-check, warning=FALSE}
pp_check(time.with(sample_prior = "only"), nsamples = 200) + scale_x_log10()
```

#### Beta Parameter Influence

```{r priors-beta, warning=FALSE}
sim.size <- 1000
sim.intercept <- rnorm(sim.size, 7.5, 1)
sim.beta <- rnorm(sim.size, 0, 0.5)
sim.beta.diff <- exp(sim.intercept + sim.beta) - exp(sim.intercept)
sim.beta.diff.min <- sim.beta.diff / 60

data.frame(x = sim.beta.diff.min) %>%
  ggplot(aes(x)) +
  geom_density() +
  xlim(-50, 50) +
  labs(
    title = "Beta parameter prior influence",
    x = "Time difference (min)",
    y = "Density"
  )

```

### Model fit {.tabset}

We check the posterior distribution and can see that the model seems to have been able to fit the data well. 
Sampling seems to also have worked well as Rhat values are close to 1 and the sampling plots look nice.

#### Posterior Predictive check

```{r base-pp-check}
pp_check(time.with(), nsamples = 200) + scale_x_log10()
```

#### Summary

```{r base-summary}
summary(time.with())
```

#### Sampling plots

```{r base-plot}
plot(time.with(), ask = FALSE)
```

## Model predictor extenstions {.tabset}

```{r mo-priors}
# default prior for monotonic predictor
edlvl_prior <- prior(dirichlet(2), class = "simo", coef = "moeducation_level1")
```

We use `loo` to check some possible extensions on the model.

### One variable {.tabset}

```{r model-extension-1, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  
  # New model(s)
  time.with("work_domain"),
  time.with("work_experience_programming.s"),
  time.with("work_experience_java.s"),
  time.with("education_field"),
  time.with("mo(education_level)", edlvl_prior),
  time.with("workplace_peer_review"),
  time.with("workplace_td_tracking"),
  time.with("workplace_pair_programming"),
  time.with("workplace_coding_standards"),
  time.with("scenario"),
  time.with("group")
)
```

#### Comparison

```{r model-extension-1-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-1-dig, warning=FALSE}
loo_result[1]
```

### Two variables {.tabset}

```{r model-extension-2, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  time.with("scenario"),
  time.with("education_field"),
  time.with("mo(education_level)", edlvl_prior),
  
  # New model(s)
  time.with(c("scenario", "education_field")),
  time.with(c("scenario", "mo(education_level)"), edlvl_prior),
  time.with(c("education_field", "mo(education_level)"), edlvl_prior)
)
```

#### Comparison

```{r model-extension-2-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-2-dig, warning=FALSE}
loo_result[1]
```


### Three variables {.tabset}

```{r model-extension-3, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  time.with("scenario"),
  time.with(c("scenario", "education_field")),
  time.with(c("scenario", "mo(education_level)"), edlvl_prior),
  
  # New model(s)
  time.with(c("scenario", "mo(education_level)", "education_field"), edlvl_prior)
)
```

#### Comparison

```{r model-extension-3-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-3-dig, warning=FALSE}
loo_result[1]
```


## Candidate models  {.tabset}
We pick some of our top performing models as candidates and inspect them closer.

The candidate models are named and listed in order of complexity.

### Time0  {.tabset}

We select the simplest model as a baseline.

```{r time0, class.source = 'fold-show'}
time0 <- brm(
  "time ~ 1 + high_debt_version + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time0-sum}
summary(time0)
```

#### Random effects

```{r time0-raneff}
ranef(time0)
```

#### Sampling plots

```{r time0-plot}
plot(time0, ask = FALSE)
```

#### Posterior predictive check

```{r time0-pp}
pp_check(time0, nsamples = 200) + scale_x_log10()
```

### Time1  {.tabset}

We select the best performing model with one variable.

```{r time1, class.source = 'fold-show'}
time1 <- brm(
  "time ~ 1 + high_debt_version + scenario + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time1",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time1-sum}
summary(time1)
```

#### Random effects

```{r time1-raneff}
ranef(time1)
```

#### Sampling plots

```{r time1-plot}
plot(time1, ask = FALSE)
```

#### Posterior predictive check

```{r time1-pp}
pp_check(time1, nsamples = 200) + scale_x_log10()
```

### Time2  {.tabset}

We select the best performing model with two variables.

```{r time2, class.source = 'fold-show'}
time2 <- brm(
  "time ~ 1 + high_debt_version + scenario + mo(education_level) + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape"),
    prior(dirichlet(2), class = "simo", coef = "moeducation_level1")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time2",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time2-sum}
summary(time2)
```

#### Random effects

```{r time2-raneff}
ranef(time2)
```

#### Sampling plots

```{r time2-plot}
plot(time2, ask = FALSE)
```

#### Posterior predictive check

```{r time2-pp}
pp_check(time2, nsamples = 200) + scale_x_log10()
```

### Time3  {.tabset}

We select the second best performing model with two variables.

```{r time3, class.source = 'fold-show'}
time3 <- brm(
  "time ~ 1 + high_debt_version + scenario + education_field + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time3",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time3-sum}
summary(time3)
```

#### Random effects

```{r time3-raneff}
ranef(time3)
```

#### Sampling plots

```{r time3-plot}
plot(time3, ask = FALSE)
```

#### Posterior predictive check

```{r time3-pp}
pp_check(time3, nsamples = 200) + scale_x_log10()
```

## Final Model 
All candidate models look nice, none is significantly better than the others, we will proceed the simplest model: `time0`

### Variations {.tabset}
We will try a few different variations of the selected candidate model.

#### All data points {.tabset}

Some participants did only complete one scenario. Those has been excluded from the initial dataset to improve sampling of the models. We do however want to use all data we can and will therefore try to fit the model with the complete dataset.

```{r variation.all, class.source = 'fold-show'}
time0.all <- brm(
  "time ~ 1 + high_debt_version + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0.all",
  file_refit = "on_change",
  seed = 20210421
)
```

##### Summary

```{r variation.all-sum}
summary(time0.all)
```

##### Random effects

```{r variation.all-raneff}
ranef(time0.all)
```

##### Sampling plots

```{r variation.all-plot}
plot(time0.all, ask = FALSE)
```

##### Posterior predictive check

```{r variation.all-pp}
pp_check(time0.all, nsamples = 200) + scale_x_log10()
```

#### With experience predictor {.tabset}

As including all data points didn't harm the model we will create this variant with all data points as well.

This variation includes `work_experience_programming.s` predictors as it can give further insight into how experience play a factor in the effect we try to measure. This is especially important as our sampling shewed towards containing less experienced developer than the population at large.

```{r variation.all.exp, class.source = 'fold-show'}
time0.all.exp <- brm(
  "time ~ 1 + high_debt_version + work_experience_programming.s + (1 | session)",
  prior = c(
    prior(normal(0, 0.5), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0.all.exp",
  file_refit = "on_change",
  seed = 20210421
)
```

##### Summary

```{r variation.all.exp-sum}
summary(time0.all.exp)
```

##### Random effects

```{r variation.all.exp-raneff}
ranef(time0.all.exp)
```

##### Loo comparison

```{r variation.all.exp-loo, warning=FALSE}
loo(
  time0.all,
  time0.all.exp
)
```

##### Sampling plots

```{r variation.all.exp-plot}
plot(time0.all.exp, ask = FALSE)
```

##### Posterior predictive check

```{r variation.all.exp-pp}
pp_check(time0.all.exp, nsamples = 200) + scale_x_log10()
```

### Final model
* Fitting the model to all data point did not significantly damage the model and will be used as is a more fair representation of reality.
* Adding the experience predictors did not significantly damage the model and will be used as it provides useful insight.

This means that our final model, with all data points and experience predictors, is `time0.all.exp`

## Interpreting the model
To begin interpreting the model we look at how it's parameters were estimated. As our research is focused on how the outcome of the model is effected we will mainly analyze the $\beta$ parameters.

### $\beta$ parameters
```{r interpret-beta-plot, warning=FALSE, message=FALSE}
mcmc_areas(time0.all.exp, pars = c("b_high_debt_versionfalse", "b_work_experience_programming.s"), prob = 0.95) + scale_y_discrete() +
  scale_y_discrete(labels=c("High debt version: false", "Professional programming experience")) +
  ggtitle("Beta parameters densities in time model", subtitle = "Shaded region marks 95% of the density. Line marks the median")
```



### Effects sizes
We start by extracting posterior samples
```{r effect-size}
scale_programming_experience <- function(x) {
  (x - mean(d.completed$work_experience_programming))/ sd(d.completed$work_experience_programming)
}
unscale_programming_experience <- function(x) {
  x * sd(d.completed$work_experience_programming) + mean(d.completed$work_experience_programming)
}

post_settings <- expand.grid(
  high_debt_version = c("false", "true"),
  session = NA,
  work_experience_programming.s = sapply(c(0, 3, 10, 25, 40), scale_programming_experience)
)

post <- posterior_predict(time0.all.exp, newdata = post_settings) %>%
  melt(value.name = "estimate", varnames = c("sample_number", "settings_id")) %>%
  left_join(
    rowid_to_column(post_settings, var= "settings_id"),
    by = "settings_id"
  ) %>%
  mutate(work_experience_programming = unscale_programming_experience(work_experience_programming.s)) %>%
  select(
    estimate,
    high_debt_version,
    work_experience_programming
  ) %>%
  mutate(estimate = estimate/60)

ggplot(post, aes(x=estimate, fill = high_debt_version)) +
  geom_density(alpha = 0.5) +
  scale_x_log10() +
  scale_fill_manual(
    name = "Debt version",
    labels = c("Low debt", "High debt"),
    values = c("lightblue", "darkblue")
  ) +
  facet_grid(rows = vars(work_experience_programming)) +
  labs(
    title = "Time to complete task / years of programming experience",
    subtitle = "Notice! x-axis is log10 scaled.",
    x = "Time (min)",
    y = "Density"
  )

```

```{r effect-size-diff, warning=FALSE, message=FALSE}
post.diff <- post %>% filter(high_debt_version == "true")
post.diff$estimate = post.diff$estimate -  filter(post, high_debt_version == "false")$estimate

ggplot(post.diff, aes(x=estimate, y = 0, fill = stat(quantile))) +
  geom_density_ridges_gradient(quantile_lines = TRUE, quantile_fun = hdi, vline_linetype = 2) +
  scale_fill_manual(name = "HDI", labels = c("100%", "95%", "100%"), values = c("transparent", "lightblue", "transparent"),) +
  xlim(-100, 100) +
  facet_grid(rows = vars(work_experience_programming)) +
  labs(
    title = "Time to complete task / years of programming experience",
    subtitle = "Difference as: high debt time - low debt time",
    x = "Time (min)",
    y = "Density"
  )
```
