---
title: "Time Outcome"
author: Hampus Broman & William Levén
date: 2021-04
output: 
  html_document: 
    pandoc_args: [ "-o", "html/time.html" ]
---

```{r include-setup, include=FALSE}
# Load setup file
source(knitr::purl('setup.Rmd', output = tempfile()))
```


## Looking at the data

We plot the data and can see that there is no obvious large difference between the versions with high and low debt

```{r plot}
d.both_completed %>%
  ggplot(aes(x=time/60, fill=high_debt_version)) + 
  geom_boxplot() +
  labs(
    title = "Distribution of time measurements for the different debt levels",
    subtitle = "Notice! Log10 x-scale",
    x ="Time (min)"
  ) +
  scale_y_continuous(breaks = NULL) +
  scale_x_log10() +
  scale_fill_discrete(name = "Debt Level", labels = c("High Debt", "Low Debt"), guide = guide_legend(reverse = TRUE)) +
  theme_minimal()
```

## Descriptive Statistics

```{r descriptive-statistics}
d.both_completed %>%
  pull(time) %>% 
  summary()

sprintf("Variance: %.0f", var(pull(d.both_completed, time)))
```

## Initial model
As the variance is much greater than the mean we will use a negative binomial family that allows us to model the variance separately.

We include `high_debt_verison` as a predictor in our model as this variable represent the very effect we want to measure.
We also include a varying intercept for each individual to prevent the model from learning too much from single participants with extreme measurements.

### Selecting Priors {.tabset}

We iterate over the model until we have sane priors.

#### Base model with priors

```{r initial-model-definition, class.source = 'fold-show'}
time.with <- extendable_model(
  base_name = "time",
  base_formula = "time ~ 1 + high_debt_version + (1 | session)",
  base_priors = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = d.both_completed,
  base_control = list(adapt_delta = 0.95)
)
```

#### Default priors

```{r default-priors}
prior_summary(time.with(only_priors= TRUE))
```

#### Selected priors

```{r selected-priors}
prior_summary(time.with(sample_prior = "only"))
```

#### Prior Predictive Check

```{r priors-check, warning=FALSE}
pp_check(time.with(sample_prior = "only"), nsamples = 200) + 
  scale_x_log10()
```

### Model fit {.tabset}

We check the posterior distribution and can see that the model seems to have been able to fit the data well. 
Sampling seems to also have worked well as Rhat values are close to 1 and the sampling plots look nice.

#### Posterior Predictive check

Notice: log10 scale on x-axis.

```{r base-pp-check}
pp_check(time.with(), nsamples = 100) +
  scale_x_log10()
```

#### Summary

```{r base-summary}
summary(time.with())
```

#### Sampling plots

```{r base-plot}
plot(time.with(), ask = FALSE)
```

## Model predictor extenstions {.tabset}

```{r mo-priors}
# default prior for monotonic predictor
edlvl_prior <- prior(dirichlet(2), class = "simo", coef = "moeducation_level1")
```


We use `loo` to check some possible extensions on the model.

### One variable {.tabset}

```{r model-extension-1, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  
  # New model(s)
  time.with("work_domain"),
  time.with("work_experience_programming.s"),
  time.with("work_experience_java.s"),
  time.with("education_field"),
  time.with("mo(education_level)", edlvl_prior),
  time.with("workplace_peer_review"),
  time.with("workplace_td_tracking"),
  time.with("workplace_pair_programming"),
  time.with("workplace_coding_standards"),
  time.with("scenario"),
  time.with("group")
)
```

#### Comparison

```{r model-extension-1-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-1-dig, warning=FALSE}
loo_result[1]
```

### Two variables {.tabset}

```{r model-extension-2, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  time.with("scenario"),
  time.with("education_field"),
  time.with("workplace_peer_review"),
  time.with("mo(education_level)", edlvl_prior),
  
  # New model(s)
  time.with(c("scenario", "education_field")),
  time.with(c("scenario", "mo(education_level)"), edlvl_prior),
  time.with(c("scenario", "workplace_peer_review")),
  time.with(c("education_field", "mo(education_level)"), edlvl_prior),
  time.with(c("education_field", "workplace_peer_review")),
  time.with(c("mo(education_level)", "workplace_peer_review"), edlvl_prior)
)
```

#### Comparison

```{r model-extension-2-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-2-dig, warning=FALSE}
loo_result[1]
```


### Three variables {.tabset}

```{r model-extension-3, warning=FALSE, class.source = 'fold-show'}
loo_result <- loo(
  # Benchmark model(s)
  time.with(),
  time.with("scenario"),
  time.with(c("scenario", "education_field")),
  time.with(c("scenario", "mo(education_level)"), edlvl_prior),
  
  # New model(s)
  time.with(c("scenario", "mo(education_level)", "education_field"), edlvl_prior)
)
```

#### Comparison

```{r model-extension-3-sum, warning=FALSE}
loo_result[2]
```

#### Diagnostics

```{r model-extension-3-dig, warning=FALSE}
loo_result[1]
```


## Candidate models  {.tabset}
We pick some of our top performing models as candidates and inspect them closer.

The candidate models are named and listed in order of complexity.

### Time0  {.tabset}

We select the simplest model as a baseline.

```{r time0, class.source = 'fold-show'}
time0 <- brm(
  "time ~ 1 + high_debt_version + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time0-sum}
summary(time0)
```

#### Random effects

```{r time0-raneff}
ranef(time0)
```

#### Sampling plots

```{r time0-plot}
plot(time0, ask = FALSE)
```

#### Posterior predictive check

```{r time0-pp}
pp_check(time0, nsamples = 150) + scale_x_log10()
```

### Time1  {.tabset}

We select the best performing model with one variable.

```{r time1, class.source = 'fold-show'}
time1 <- brm(
  "time ~ 1 + high_debt_version + scenario + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time1",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time1-sum}
summary(time1)
```

#### Random effects

```{r time1-raneff}
ranef(time1)
```

#### Sampling plots

```{r time1-plot}
plot(time1, ask = FALSE)
```

#### Posterior predictive check

```{r time1-pp}
pp_check(time1, nsamples = 150) + scale_x_log10()
```

### Time2  {.tabset}

We select the best performing model with two variables.

```{r time2, class.source = 'fold-show'}
time2 <- brm(
  "time ~ 1 + high_debt_version + scenario + mo(education_level) + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape"),
    prior(dirichlet(2), class = "simo", coef = "moeducation_level1")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time2",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time2-sum}
summary(time2)
```

#### Random effects

```{r time2-raneff}
ranef(time2)
```

#### Sampling plots

```{r time2-plot}
plot(time2, ask = FALSE)
```

#### Posterior predictive check

```{r time2-pp}
pp_check(time2, nsamples = 150) + scale_x_log10()
```

### Time3  {.tabset}

We select the second best performing model with two variables.

```{r time3, class.source = 'fold-show'}
time3 <- brm(
  "time ~ 1 + high_debt_version + scenario + education_field + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.both_completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time3",
  file_refit = "on_change",
  seed = 20210421
)
```

#### Summary

```{r time3-sum}
summary(time3)
```

#### Random effects

```{r time3-raneff}
ranef(time3)
```

#### Sampling plots

```{r time3-plot}
plot(time3, ask = FALSE)
```

#### Posterior predictive check

```{r time3-pp}
pp_check(time3, nsamples = 150) + scale_x_log10()
```

## Final Model 
All candidate models look nice, none is significantly better than the others, we will proceed the simplest model: `time0`

### Variations {.tabset}
We will try a few different variations of the selected candidate model.

#### All data points {.tabset}

Some participants did only complete one scenario. Those has been excluded from the initial dataset to improve sampling of the models. We do however want to use all data we can and will therefore try to fit the model with the complete dataset.

```{r variation.all, class.source = 'fold-show'}
time0.all <- brm(
  "time ~ 1 + high_debt_version + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0.all",
  file_refit = "on_change",
  seed = 20210421
)
```

##### Summary

```{r variation.all-sum}
summary(time0.all)
```

##### Random effects

```{r variation.all-raneff}
ranef(time0.all)
```

##### Sampling plots

```{r variation.all-plot}
plot(time0.all, ask = FALSE)
```

##### Posterior predictive check

```{r variation.all-pp}
pp_check(time0.all, nsamples = 150) + scale_x_log10()
```

#### With experience predictor {.tabset}

As including all data points didn't harm the model we will create this variant with all data points as well.

```{r variation.all.exp, class.source = 'fold-show'}
time0.all.exp <- brm(
  "time ~ 1 + high_debt_version + work_experience_programming.s + (1 | session)",
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(7.5, 1), class = "Intercept"),
    prior(exponential(1), class = "sd"),
    prior(gamma(0.01, 0.01), class = "shape")
  ),
  family = negbinomial(),
  data = as.data.frame(d.completed),
  control = list(adapt_delta = 0.95),
  file = "fits/time0.all.exp",
  file_refit = "on_change",
  seed = 20210421
)
```

##### Summary

```{r variation.all.exp-sum}
summary(time0.all.exp)
```

##### Random effects

```{r variation.all.exp-raneff}
ranef(time0.all.exp)
```

##### Loo comparison

```{r variation.all.exp-loo, warning=FALSE}
loo(
  time0.all,
  time0.all.exp
)
```

##### Sampling plots

```{r variation.all.exp-plot}
plot(time0.all.exp, ask = FALSE)
```

##### Posterior predictive check

```{r variation.all.exp-pp}
pp_check(time0.all.exp, nsamples = 150) + scale_x_log10()
```

### Final model
* Fitting the model to all data point did not significantly damage the model and will be used as is a more fair representation of reality.
* Adding the experience predictors did not significantly damage the model and will be used as it provides useful insight.

This means that our final model, with all data points and experience predictors, is `time0.all.exp`

## Interpreting the model

```{r}
summary(time0.all.exp, prob = 0.89)
```

### Effects sizes
We start by extracting posterior samples
```{r}
post <- posterior_predict(time0.all.exp, newdata = data.frame(
  high_debt_version = c("false", "true"),
  session = NA,
  work_experience_programming.s = (10 - mean(d.completed$work_experience_programming))/ sd(d.completed$work_experience_programming)
))
post.low <-  post[,1]
post.high <- post[,2]
summary(data.frame(
  "Low Debt"=post.low,
  "High Debt"=post.high
))
post.diff <- post.high - post.low
post.diff.scaled <- post.diff / mean(post)
```


When we look at the distribution of the data between the high and low debt version we can see a difference, they are however quite similar.
```{r}
ggplot(data.frame(time = post.high)) +
  geom_density(data = data.frame(time = post.low), aes(x=time/60, colour = "blue")) +
  geom_density(data = data.frame(time = post.high),  aes(x=time/60, colour = "red", linetype="dashed" )) +
  scale_x_log10() +
  labs(
    title = "Posterior density of time for high and low debt versions",
    subtitle = "Notice! x-axis is log10 scaled.",
    x ="Time (min)",
    y = "Density"
  ) +
  theme_minimal() +
  scale_y_continuous(breaks = NULL) +
  scale_colour_manual(name="Version",labels=c("Low Debt", "High Debt"), values=c("blue", "red"))

```

We can also plot the difference between time for the high debt version and the low debt version. In those graphs we see that the difference is centered around zero with slightly more weight on the positive (indicating slightly longer times for the high debt version)
```{r}

ggplot(data.frame(x = post.diff/60)) + geom_density(aes(x=x)) +
  labs(
    title = "Posterior density of time difference low and high debt versions ",
    subtitle = "Difference = High Debt time - Low Debt time",
    x ="Time Difference (min)",
    y = "Density"
  ) +
  theme_minimal() +
  scale_y_continuous(breaks = NULL)


data.frame(x = post.diff.scaled) %>%
  ggplot() +
  geom_density(aes(x=x)) +
  labs(
    title = "Posterior density of time difference low and high debt versions ",
    subtitle = "Difference = High Debt time - Low Debt time",
    x ="Time Difference (prortion of mean time)",
    y = "Density"
  ) +
  theme_minimal() +
  scale_y_continuous(breaks = NULL)

```


We can also check the probability of the high debt effect being less than as well as grather then zero.
```{r}
sprintf("high debt effect < 0: %.2f%%", sum(sign(post.diff) == -1)/length(post.diff) * 100)
sprintf("high debt effect > 0: %.2f%%", sum(sign(post.diff) == 1)/length(post.diff) * 100)
```

We can see no "significant" relation between the amount of technical debt in the given codebase and time spent on the task.

